---
title: "Reduce mental burden in plots"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: true
    toc: yes
    toc_float: yes
---



```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r libraries, include=FALSE}
library(tidyverse)
library(extrafont)
library(vroom)
library(ggtext)
library(gapminder)
library(ggrepel)
library(patchwork)
library(gghighlight)
library(skimr)

# remotes::install_github("kjhealy/nycdogs")
library(nycdogs)
library(sf) # for geospatial visualisation

# loadfonts(device="pdf") if you have a mac
# loadfonts(device="win") if you have a windows machine

```

# Who killed the emperors?

1. Read in the emperors data (no need to change this part of the code)
2. Sort the data by the number of each type of killer
3. Take a look at the data up until this point. Pick something you find interesting that you want to highlight. Then, in `mutate()`, create a new variable that is `TRUE` if `killer` matches the category you want to highlight and `FALSE` otherwise
4. Use the variable you just created in the `fill` aesthetic of the ggplot call
5. Finally, use `scale_fill_manual()` to add the fill colours grey70 = #B3B3B3 for the background, and tomato = #FF6347 to higlights. 


```{r, emperors, eval=FALSE}
# make sure you remove eval=FALSE before knitting the document

# source of data https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-08-13
url <- "https://github.com/rfordatascience/tidytuesday/raw/master/data/2019/2019-08-13/emperors.csv"
emperors <- vroom(url)


emperor_killers <- emperors %>%
  count(killer) 

#if you want to use hex codes, gplots::col2hex("colour name") will give you hex code
# grey70 = #B3B3B3, tomato = #FF6347

# Besides 'tomato' experiment with a different colour to highlight
# https://color.adobe.com/create/color-wheel
# https://www.color-hex.com/ 

#define colours to use: grey70 for everything, tomato for assassination
my_colours <- c("grey70", "tomato")
my colours <- c("#b3b3b3", "#ff6347")


emperor_killers %>% 
  mutate(
    ___________ = ifelse(killer == "___________", TRUE, FALSE),
    # use `fct_inorder()` to maintain the way we sorted the data
    killer = fct_inorder(killer)
  ) %>%
  ggplot(aes(x = n, y = killer, fill = ___________)) +
  geom_col() +
  scale_fill_manual(values = my_colours)+
  geom_text(
    aes(label = n, x = n - .25), 
    color = "white", 
    size = 5,
    hjust = 1
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_blank(), 
    legend.position = "none" 
  ) +
  ___________(values = ___________) +
  xlab("number of emperors")
```


# Add legends for just some gapminder countries

1. Use `sample()` to select 20 random countries from the Gapminder dataset to plot (run the `set.seed()` line first if you want the same results)
2. In the `mutate()` call, check if `country` is one of the countries in `some_countries`. If it's not, make the label an empty string (""),
3. Add the text repel geom from ggrepel. Set the `label` aesthetic using the variable just created in `mutate()`

```{r, gapminder_some_countries, eval = FALSE}
# make sure you remove eval=FALSE before knitting the document

set.seed(42)

some_countries <- gapminder$country %>% 
  levels() %>% 
  sample(20) 

# what if we try to add label for all countries?
gapminder %>%
  filter(year == 2007) %>%
  ggplot(aes(log(gdpPercap), lifeExp)) +
    geom_point(
    size = 3,
    alpha = 0.8,
    shape = 21,
    colour = "white",
    fill = "#001e62"
  )+
  theme_minimal() +
  geom_text_repel(aes(label=country))+
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Gapminder 2007: Life expectancy vs GDP per capita",
    x = "log(GDP per capita)",
    y = "Life Expectancy"
  )+
  NULL

# let us only add labels to the 20 randomly chosen countries
gapminder %>%
  filter(year == 2007) %>%
  mutate(
    label = ifelse(______  %in% _____, as.character(country), _____________ )
  ) %>%
  ggplot(aes(log(gdpPercap), lifeExp)) +
    geom_point(
    size = 3,
    alpha = 0.8,
    shape = 21,
    colour = "white",
    fill = "#001e62"
  )+
  theme_minimal() +
  geom_text_repel(aes(label = label))+
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Gapminder 2007: Life expectancy vs GDP per capita",
    x = "log(GDP per capita)",
    y = "Life Expectancy"
  )
```


# Were there more men or women on the Titanic?

```{r, titanic}

# adapted from https://clauswilke.com/dataviz/histograms-density-plots.html#multiple-histograms-densities

# load titanic data
titanic <- vroom::vroom(here::here("data", "titanic3.csv"))

# First, explore effect of different values of bandwith in geom_density()
# bw	:The smoothing bandwidth to be used, siilar to a histogram's binwidth

pdens1 <- ggplot(titanic, aes(x = age)) + 
  geom_density(fill = "#56B4E9", color = "#56B4E9", bw = .5, kernel = "gaussian") + 
  scale_y_continuous(limits = c(0, 0.046), expand = c(0, 0), name = "density") +
  scale_x_continuous(name = "age (years)", limits = c(0, 75), expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  theme_minimal(12) +
  theme(
    axis.line.x = element_blank(),
    plot.margin = margin(3, 1.5, 3, 1.5)
  )

pdens2 <- ggplot(titanic, aes(x = age)) + 
 geom_density(fill = "#56B4E9", color = "#56B4E9", bw = 2, kernel = "gaussian") + 
  scale_y_continuous(limits = c(0, 0.046), expand = c(0, 0), name = "density") +
  scale_x_continuous(name = "age (years)", limits = c(0, 75), expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  theme_minimal(12)  +
  theme(
    axis.line.x = element_blank(),
    plot.margin = margin(3, 1.5, 3, 1.5)
  )

pdens3 <- ggplot(titanic, aes(x = age)) + 
 geom_density(fill = "#56B4E9", color = "#56B4E9", bw = 5, kernel = "gaussian") + 
  scale_y_continuous(limits = c(0, 0.046), expand = c(0, 0), name = "density") +
  scale_x_continuous(name = "age (years)", limits = c(0, 75), expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  theme_minimal(12)  +
  theme(
    axis.line.x = element_blank(),
    plot.margin = margin(3, 1.5, 3, 1.5)
  )

pdens4 <- ggplot(titanic, aes(x = age)) + 
  geom_density(fill = "#56B4E9", color = "#56B4E9", bw = 2, kernel = "rectangular") + 
  scale_y_continuous(limits = c(0, 0.046), expand = c(0, 0), name = "density") +
  scale_x_continuous(name = "age (years)", limits = c(0, 75), expand = c(0, 0)) +
  coord_cartesian(clip = "off") +
  theme_minimal(12)  +
  theme(
    axis.line.x = element_blank(),
    plot.margin = margin(3, 1.5, 3, 1.5)
  )

# use patchwork to arrange plots
(pdens1 + pdens2) / (pdens3 + pdens4) 


# Define a vector of colours... grey80 for background, and two different colours for male/female
density_colours <- c(
  "male" = "#0072B2",
  "female" = "#D55E00",
  "all passengers" = "grey80"
)

# the default density plot + facet_wrap-- difference between men and women seems small
titanic %>%
  drop_na(sex, age) %>% # remove NAs for sex and age
  ggplot(aes(age)) +
  geom_density(aes(fill = sex, colour = sex), alpha = 0.7) +
  scale_color_manual(name = NULL, values = density_colours) +
  scale_fill_manual(name = NULL, values = density_colours) +
  theme_minimal(16) +
  theme(legend.position = "bottom", legend.justification = "center")+
  facet_wrap(~sex)+
  NULL




titanic %>%
  drop_na(sex, age) %>% # remove NAs for sex and age, and pass dataframe to ggplot
  
  # in the ggplot(), besides `x = age`, add `y = ..count..` to `aes()`
  ggplot(aes(x = age, y = ..count..)) + 
  
  
  # Add an additional `geom_density()` that will draw the background density, for all participants. 
  # This should come *before* the 'normal' `geom_density()` so that it draws the background.
  geom_density(
    
    # In the background geom_density(), set the `data` argument to be a function. This function 
    # takes a data frame and removes sex (select x, -sex) (on which sex we will facet on later).
    data = function(x) select(x, -sex),
    
    # Set both `colour` and `fill` equal to "all passengers", *not* `sex`.
    aes(fill = "all passengers", colour = "all passengers")
  ) +
  
  # now that we have the grey background, we can plot densities
  # coloured and facet_wrapped by sex
  geom_density(aes(fill = sex, colour = sex), alpha = 0.7, bw = 2) +
  
  # `facet_wrap()` to facet the plot by `sex`.
  # labeller() makes it easy to assign different labellers to different factors
  facet_wrap(~sex, labeller = labeller(sex = function(sex) paste(sex, "passengers"))) +
  
  
  # use the colours you defined for male, female, and all participants
  scale_colour_manual(
    values = density_colours,
    breaks = c("male", "female", "all passengers"),
    labels = c( "males", "females","all passengers"),
    name = NULL,
    guide = guide_legend(direction = "horizontal")
  )+
  scale_fill_manual(
    values = density_colours,
    breaks = c("male", "female", "all passengers"),
    labels = c( "males", "females","all passengers"),
    name = NULL,
    guide = guide_legend(direction = "horizontal")
  )+
  scale_x_continuous(name = "age (years)", limits = c(0, 75), expand = c(0, 0)) +
  scale_y_continuous(limits = c(0, 40), expand = c(0, 0), name = "scaled density") +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        legend.justification = "center",
        text=element_text(size=12, family="Lato"),
        plot.title.position = "plot"
  ) +
  NULL

  
```

# `gghighlight`

We want to create a graph that looks at African countriews were life expectancy between 1992 and 2007 went down.

```{r}
africa <- gapminder %>%
  filter(continent == "Africa") %>% 
  # sort by life expectancy in 1952, the first year in the dataframe
  arrange(year, desc(lifeExp)) %>% 
  mutate(country = fct_inorder(factor(country)))

# find the African countries that had better life expectancy in 1992 compared to 2007
life_expectancy_dropped <- africa %>%
  
  # pivot wider to be able to compare life expectancy in 1992 to 2007,
  pivot_wider(country, names_from = year, values_from = lifeExp) %>%
  
  # create new variables, `le_dropped`, that is `TRUE` if life expectancy was higher in 1992
  # and `le_delta` which is simply the difference between life expectancy in 2007 and 1992
  mutate(
    le_delta = `2007` - `1992`,
    le_dropped = `1992` > `2007`) %>% 
  select(-c(2:13))

# join `le_dropped` to each observation for each country
africa <- left_join(africa, life_expectancy_dropped, by = "country")

le_line_plot <- africa %>%
  ggplot(aes(year, lifeExp, color = country, group = country)) +
  geom_line(size = 1.2, alpha = 0.9, color = "#E58C23") +
  theme_minimal(base_size = 14) +
  
  # let us add a vertical line at 1992, when we start our comparison
  geom_vline(xintercept = 1992, alpha = 0.1)+

  theme(
    # remove legend using the legend.position argument in theme()
    # ________________,
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank()
  ) +
  labs(
    title = "Life expectancy reduction, 1992 to 2007",
    y = "life expectancy", 
    caption = "sorted by life expectancy in 1952") +
  scale_x_continuous(breaks = seq(1952, 2007, 10))+
  NULL

le_line_plot


# Using `gghighlight()` we add add direct labels to the plot. 
# The first argument defines which lines to highlight using `le_dropped`. 
# Also add the arguments `use_group_by = FALSE` and `unhighlighted_colour = "grey80"`. 

le_line_plot + 
  gghighlight(
    le_dropped,
    use_group_by = FALSE,
    unhighlighted_colour = "grey80"
  )

#  If we facet_wrap() we do not need direct labels, so set `use_direct_label = FALSE` 

le_line_plot +
  gghighlight(
    le_dropped,
    use_group_by = FALSE,
    use_direct_label = FALSE,
    unhighlighted_colour = "grey80"
  ) +
  facet_wrap(~country)
```

# On your own: NYC Dogs 

This is to get you started on your own. You can install the relevant `nycdogs` package and read more about it here https://kjhealy.github.io/nycdogs/ 

```{r, nyc_dogs}

# You can install nycdogs from GitHub with:
# devtools::install_github("kjhealy/nycdogs")

# load the data
library(nycdogs)
data("nyc_license")
data("nyc_zips")

# what is the prevalent NYC breed ?
nyc_license %>% 
  count(breed_rc, sort=TRUE) 

# what is the prevalent NYC dog name by animal gender? 
# Bella for girls, Max for boys!

nyc_license %>% 
  count(animal_name, animal_gender, sort=TRUE) 



# Maps for two Breeds
top_breeds <- nyc_license %>%
  group_by(zip_code, breed_rc) %>%
  tally() %>%
  filter(n>10) %>% 
  mutate(freq = n / sum(n),
         pct = round(freq*100, 2)) %>%
  filter(breed_rc == "Yorkshire Terrier" |  breed_rc == "Labrador (or Crossbreed)") 



top_breeds_map <- left_join(nyc_zips, top_breeds) %>% 
  na.omit() 


top_breeds_map %>% ggplot(mapping = aes(fill = pct)) +
  geom_sf(color = "gray80", size = 0.1) +
  scale_fill_viridis_b(option = "A", direction= -1) +
  labs(fill = "Percent of Licensed Dogs") +
  facet_wrap(~breed_rc)+
  theme_void() + 
  guides(fill = guide_legend(title.position = "top", 
                             label.position = "bottom")) 


```
# On your own: Coffee ratings

Coffee ratings. Which country produces the best coffee? Does location and altitude of cultivation play a role?

You can see someone's analysis here https://towardsdatascience.com/the-data-speak-ethiopia-has-the-best-coffee-91f88ed37e84



```{r, coffee_ratings}

url <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv'

coffee_ratings <- vroom::vroom(url)

skim(coffee_ratings)

```


