---
title: "Homework 4: Machine Learning"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false
options(scipen = 999) #disable scientific notation
library(tidyverse)
library(tidymodels)
library(GGally)
library(sf)
library(leaflet)
library(janitor)
library(rpart.plot)
library(here)
library(scales)
library(vip)
```


# The Bechdel Test

https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/

The [Bechdel test](https://bechdeltest.com) is a way to assess how women are depicted in Hollywood movies.  In order for a movie to pass the test:

1. It has to have at least two [named] women in it
2. Who talk to each other
3. About something besides a man

There is a nice article and analysis you can find here https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/
We have a sample of 1394 movies and we want to fit a model to predict whether a film passes the test or not.

```{r read_data}

bechdel <- read_csv(here::here("data", "bechdel.csv")) %>% 
  mutate(test = factor(test)) 
glimpse(bechdel)

```
How many films fail/pass the test, both as a number and as a %?

```{r}

```


## Movie scores
```{r}
ggplot(data = bechdel, aes(
  x = metascore,
  y = imdb_rating,
  colour = test
)) +
  geom_point(alpha = .3, size = 3) +
  scale_colour_manual(values = c("tomato", "olivedrab")) +
  labs(
    x = "Metacritic score",
    y = "IMDB rating",
    colour = "Bechdel test"
  ) +
 theme_light()
```


# Split the data
```{r}
# **Split the data**

set.seed(123)

data_split <- initial_split(bechdel, # updated data
                           prop = 0.8, 
                           strata = test)

bechdel_train <- training(data_split) 
bechdel_test <- testing(data_split)
```

Check the counts and % (proportions) of the `test` variable in each set.
```{r}

```

## Feature exploration

## Any outliers? 

```{r}

bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore) %>% 

    pivot_longer(cols = 2:6,
               names_to = "feature",
               values_to = "value") %>% 
  ggplot()+
  aes(x=test, y = value, fill = test)+
  coord_flip()+
  geom_boxplot()+
  facet_wrap(~feature, scales = "free")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x=NULL,y = NULL)

```

## Scatterplot - Correlation Matrix

Write a paragraph discussing the output of the following 
```{r, warning=FALSE, message=FALSE}
bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore)%>% 
  ggpairs(aes(colour=test), alpha=0.2)+
  theme_bw()
```


## Categorical variables

Write a paragraph discussing the output of the following 
```{r}
bechdel %>% 
  group_by(genre, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
  
 
bechdel %>% 
  group_by(rated, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
```

# Train first models. `test ~ metascore + imdb_rating`

```{r}
lr_mod <- logistic_reg() %>% 
  set_engine(engine = "glm") %>% 
  set_mode("classification")

lr_mod


tree_mod <- decision_tree() %>% 
  set_engine(engine = "C5.0") %>% 
  set_mode("classification")

tree_mod 
```

```{r}


lr_fit <- lr_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )

tree_fit <- tree_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )
```

## Logistic regression

```{r}
lr_fit %>%
  broom::tidy()

lr_preds <- lr_fit %>%
  augment(new_data = bechdel_train) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0))

```

### Confusion matrix

```{r}
lr_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")


```


## Decision Tree
```{r}
tree_preds <- tree_fit %>%
  augment(new_data = bechdel) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0)) 


```

```{r}
tree_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")
```

## Draw the decision tree

```{r}
draw_tree <- 
    rpart::rpart(
        test ~ metascore + imdb_rating,
        data = bechdel_train, # uses data that contains both birth weight and `low`
        control = rpart::rpart.control(maxdepth = 5, cp = 0, minsplit = 10)
    ) %>% 
    partykit::as.party()
plot(draw_tree)

```

# Cross Validation

Run the code below. What does it return?

```{r}
set.seed(123)
bechdel_folds <- vfold_cv(data = bechdel_train, 
                          v = 10, 
                          strata = test)
bechdel_folds
```

## `fit_resamples()`

Trains and tests a resampled model.

```{r}
lr_fit <- lr_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )


tree_fit <- tree_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )
```


## `collect_metrics()`

Unnest the metrics column from a tidymodels `fit_resamples()`
```{r}

collect_metrics(lr_fit)
collect_metrics(tree_fit)


```


```{r}
tree_preds <- tree_mod %>% 
  fit_resamples(
    test ~ metascore + imdb_rating, 
    resamples = bechdel_folds,
    control = control_resamples(save_pred = TRUE) #<<
  )

# What does the data for ROC look like?
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail)  

# Draw the ROC
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail) %>% 
  autoplot()

```


# Build a better training set with `recipes`

## Preprocessing options

- Encode categorical predictors
- Center and scale variables
- Handle class imbalance
- Impute missing data
- Perform dimensionality reduction 
- ... ...

## To build a recipe

1. Start the `recipe()`
1. Define the variables involved
1. Describe **prep**rocessing [step-by-step]

## Collapse Some Categorical Levels

Do we have any `genre` with few observations?  Assign genres that have less than 3% to a new category 'Other'


```{r}
#| echo = FALSE
bechdel %>% 
  count(genre) %>% 
  mutate(genre = fct_reorder(genre, n)) %>% 
  ggplot(aes(x = genre, 
             y = n)) +
  geom_col(alpha = .8) +
  coord_flip() +
  labs(x = NULL) +
  geom_hline(yintercept = (nrow(bechdel_train)*.03), lty = 3)+
  theme_light()
```


```{r}
movie_rec <-
  recipe(test ~ .,
         data = bechdel_train) %>%
  
  # Genres with less than 5% will be in a catewgory 'Other'
    step_other(genre, threshold = .03) 
```
  

## Before recipe

```{r}
#| echo = FALSE
bechdel_train %>% 
  count(genre, sort = TRUE)
```


## After recipe

```{r}
movie_rec %>% 
  prep() %>% 
  bake(new_data = bechdel_train) %>% 
  count(genre, sort = TRUE)
```

## `step_dummy()`

Converts nominal data into numeric dummy variables

```{r}
#| results = "hide"
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_dummy(all_nominal_predictors()) 

movie_rec 
```

## Let's think about the modelling 

What if there were no films with `rated` NC-17 in the training data?

-  Will the model have a coefficient for `rated` NC-17?
-  What will happen if the test data includes a film with `rated` NC-17?

## `step_novel()`

Adds a catch-all level to a factor for any new values not encountered in model training, which lets R intelligently predict new levels in the test set.

```{r}

movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal_predictors) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal_predictors()) 

```


## `step_zv()`

Intelligently handles zero variance variables (variables that contain only a single value)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes()) 
  
```


## `step_normalize()`

Centers then scales numeric variable (mean = 0, sd = 1)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) 

```


## `step_corr()`

Removes highly correlated variables

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) %>% 
  step_corr(all_predictors(), threshold = 0.75, method = "spearman") 



movie_rec
```


# Define different models to fit

```{r}
## Model Building

# 1. Pick a `model type`
# 2. set the `engine`
# 3. Set the `mode`: regression or classification

# Logistic regression
log_spec <-  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec

# Decision Tree
tree_spec <- decision_tree() %>%
  set_engine(engine = "C5.0") %>%
  set_mode("classification")

tree_spec

# Random Forest
library(ranger)

rf_spec <- 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")


# Boosted tree (XGBoost)
library(xgboost)

xgb_spec <- 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 

# K-nearest neighbour (k-NN)
knn_spec <- 
  nearest_neighbor(neighbors = 4) %>% # we can adjust the number of neighbors 
  set_engine("kknn") %>% 
  set_mode("classification") 
```


# Bundle recipe and model with `workflows`


```{r}
log_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(movie_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

# show object
log_wflow


## A few more workflows

tree_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(tree_spec) 

rf_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(rf_spec) 

xgb_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(xgb_spec)

knn_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(knn_spec)

```

HEADS UP

1. How many models have you specified?
2. What's the difference between a model specification and a workflow?
3. Do you need to add a formula (e.g., `test ~ .`)  if you have a recipe?


# Model Comparison

You now have all your models. Adapt the code from slides `code-from-slides-CA-housing.R`, line 400 onwards to assess which model gives you the best classification. 


```{r}

```



# Deliverables

There is a lot of explanatory text, comments, etc. You do not need these, so delete them and produce a stand-alone document that you could share with someone. Knit the edited and completed R Markdown (Rmd) file as a Word or HTML document (use the "Knit" button at the top of the script editor window) and upload it to Canvas. You must be commiting and pushing your changes to your own Github repo as you go along.

# Details

-   Who did you collaborate with: TYPE NAMES HERE
-   Approximately how much time did you spend on this problem set: ANSWER HERE
-   What, if anything, gave you the most trouble: ANSWER HERE

**Please seek out help when you need it,** and remember the [15-minute rule](https://dsb2023.netlify.app/syllabus/#the-15-minute-rule){target="_blank"}. You know enough R (and have enough examples of code from class and your readings) to be able to do this. If you get stuck, ask for help from others, post a question on Slack-- and remember that I am here to help too!

> As a true test to yourself, do you understand the code you submitted and are you able to explain it to someone else?

# Rubric

13/13: Problem set is 100% completed. Every question was attempted and answered, and most answers are correct. Code is well-documented (both self-documented and with additional comments as necessary). Used tidyverse, instead of base R. Graphs and tables are properly labelled. Analysis is clear and easy to follow, either because graphs are labeled clearly or you've written additional text to describe how you interpret the output. Multiple Github commits. Work is exceptional. I will not assign these often.

8/13: Problem set is 60--80% complete and most answers are correct. This is the expected level of performance. Solid effort. Hits all the elements. No clear mistakes. Easy to follow (both the code and the output). A few Github commits.

5/13: Problem set is less than 60% complete and/or most answers are incorrect. This indicates that you need to improve next time. I will hopefully not assign these often. Displays minimal effort. Doesn't complete all components. Code is poorly written and not documented. Uses the same type of plot for each graph, or doesn't use plots appropriate for the variables being analyzed. No Github commits.
