---
title: "E628 Pre-course assignment"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
editor: visual
---

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false

library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(janitor)
library(nycflights13)
```

The goal is to test your software installation, to demonstrate competency in Markdown, and in the basics of `ggplot`.

# Install R and RStudio

You should successfully install R and R studio in your computer. We will do all of our work in this class with the open source (and free!) programming language [R](https://cran.r-project.org/). However, we will use [RStudio](https://www.rstudio.com/) as the main program to access R.

You can find details on how to [install both R and R studio here](https://am01-sep23.netlify.app/reference/01-reference/)

## Install `tidyverse`, `janitor` and `nycflights13` packages

A clean installation of R is known as **base R**. We need to install a collection of packages named `tidyverse`. Go to the packages panel in the bottom right of RStudio, click on "Install," type `tidyverse`,`janitor`, `nycflights13`  and press enter. You'll see a bunch of output in the RStudio console as all the packages are installed and, depending on the speed of your computer, this may take a while. You can also just paste and run the following

-   `install.packages("tidyverse")`
-   `install.packages("janitor")`
-   `install.packages("nycflights13")`

in the `console` (bottom left in RStudio) instead of using the packages panel.

You can find details on [installing the tidyverse here](https://am01-sep23.netlify.app/reference/02-reference/)

## Practice using Markdown

Assignments will be submitted using `Markdown` is a lightweight text formatting language that easily converts between file formats. It is integrated directly into [Quarto Markdown](https://quarto.org/docs/get-started/hello/rstudio.html), which combines R code, output, and written text into a single document (`.qmd`). You can [find more about markdown here](https://am01-sep23.netlify.app/reference/03-reference/).

Quarto allows you to use a `Visual` editor, much like using word-processing software, but you can always switch back to `Source`.

## Pandoc

[Pandoc](http://pandoc.org) is a program that converts Markdown files into basically anything else. It was created by [John MacFarlane](https://johnmacfarlane.net), a philosophy professor at the University of California, Berkeley and is widely used as a writing tool and as a basis for publishing workflow. Kieran Healy's [Plain Text Social Science workflow](http://plain-text.co) describes how to use Markdown and then convert your Markdown document to HTML, PDF, word, etc.

You should change the file name from `your-name-pre-course.Rmd` to your own name; if I were submitting, my file would be called `Kostis_Christodoulou.Rmd`

# Task 1: Short bio written using markdown

You should write within this `qmd` file a brief biography of yourself using markdown syntax. I know you have already achieved a lot, but a couple of paragraphs is more than enough.

To achieve full marks, you should include at least 4 of the following elements:

-   Headers
-   Emphasis (italics or bold)
-   Lists
-   Links
-   Embedding images

> Please delete all the intro text I wrote and start writing your short biography after this blockquote.

# Task 2: The `nycflights13` package

The `nycflights13` package contains all US domestic flights that departed out of three New York airports (JFK,Newark, La Guardia) in 2013. The package is made up of five tables which are as follows:

-   `airlines` Airline names.
-   `airports` Airport metadata
-   `flights` Flights data
-   `planes` Plane metadata.
-   `weather` Hourly weather data

## Problem 1: Use logical operators and the `flights` table to find flights that:

-   Had an arrival delay of two or more hours (\> 120 minutes)
-   Flew to Houston (IAH or HOU)
-   Were operated by United (`UA`), American (`AA`), or Delta (`DL`)
-   Departed in summer (July, August, and September)
-   Arrived more than two hours late, but didn't leave late

```{r}
#| label: problem-1

glimpse(flights)

# Had an arrival delay of two or more hours (> 120 minutes)
flights %>% 
  filter(arr_delay > 120)

# Flew to Houston (IAH or HOU)


# Were operated by United (`UA`), American (`AA`), or Delta (`DL`)


# Departed in summer (July, August, and September)


# Arrived more than two hours late, but didn't leave late

  

```

## Problem 2: The `nycflights13` package includes a table (`weather`) that describes the weather during 2013. Use that table to answer the following questions:

-   What is the distribution of temperature (`temp`) in July 2013?
-   Identify any important outliers in terms of the `wind_speed` variable.
-   What is the relationship between `dewp` and `humid`?
-   What is the relationship between `precip` and `visib`?

```{r}
#| label: problem-2

glimpse(weather)

#   What is the distribution of temperature (`temp`) in July 2013? 


#   Identify any important outliers in terms of the `wind_speed` variable.


#   What is the relationship between `dewp` and `humid`? Use a scatter plot geom_point()


#   What is the relationship between `precip` and `visib`?
  

```

## Problem 3: Use the `flights` table to answer the following questions:

-   What is the median `arr_delay` on a month-by-month basis in each `origin` airport?
-   For each `carrier`, plot the median `arr_delay` (arrival delay) for each `month` and `origin` airport.

```{r}
#| label: problem-3

glimpse(flights)

#  What is the median `arr_delay` on a month-by-month basis in each `origin` airport?
  
  
#  For each `carrier`, plot the median arrival delay for each month and `origin` airport.
  

```

# Submit the assignment

Render the completed R Markdown file as an HTML document (use the "Knit" button at the top of the script editor window) and email me your work.

## Details

If you want to, please answer the following

-   Who did you collaborate with: TYPE NAMES HERE
-   Approximately how much time did you spend on this problem set: ANSWER HERE
-   What, if anything, gave you the most trouble: ANSWER HERE
