---
title: "Homework 1"
author: "Selected answers"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```




# Data Manipulation

```{r}
library(tidyverse)
library(nycflights13)
library(showtext)

font_add_google("Montserrat") # the closest to LBS font-- can get more fonts from fonts.google.com
showtext_auto(enable = TRUE)

```



# months with  highest and lowest proportion of cancelled flights

```{r}
flights %>%
  mutate(cancelled = ifelse(is.na(dep_time), TRUE, FALSE)) %>% 
  group_by(month, cancelled) %>% 
  summarise(count = n()) %>% 
  mutate(prop = count / sum(count))

# does it matter how we group_by()? Of course it does!

flights %>%
  mutate(cancelled = ifelse(is.na(dep_time), TRUE, FALSE)) %>% 
  group_by(cancelled, month) %>% 
  summarise(count = n()) %>% 
  mutate(prop = count / sum(count))  
```

Plot cancellation % by month

```{r}

flights %>%
  # find flights that were cancelled, ie, those with no departure time
  mutate(cancelled = ifelse(is.na(dep_time), TRUE, FALSE)) %>% 
  
  # count how many cancellations/normal flights we had per month
  group_by(month, cancelled) %>% 
  summarise(count = n()) %>% 
  
  # use mutate to create a new column that calculates %
  mutate(prop = count / sum(count)) %>%
  
  # only show those flights that were cancelled
  filter(cancelled == TRUE) %>%
  
  # pipe the resulting dataframe to ggpplot
  ggplot() +
  
  # add global aesthetics
  aes(x=factor(month), y = prop)+
  
  #just use a bar chart
  geom_col()+
  
  # change the theme, to theme_light()
  theme_light()+
  
  # format y-axis labels as, e.g., 15% and not 0.15
  scale_y_continuous(labels = scales::percent)


# how about we do it by airport?

flights %>%
  mutate(cancelled = ifelse(is.na(dep_time), TRUE, FALSE)) %>% 
  group_by(origin, month, cancelled) %>% 
  summarise(count = n()) %>% 
  mutate(prop = count / sum(count)) %>% 
  filter(cancelled == TRUE) %>% 
  ggplot() +
  aes(x=factor(month), y = prop)+
  geom_col()+
  theme_light()+
  scale_y_continuous(labels = scales::percent)+
  facet_wrap(~origin, ncol = 1)

```






# Problem 3: What plane (specified by tailnum variable) traveled the most times from NYC airports in 2013? 

Please left_join() the resulting table with the table planes (also included in the nycflights13 package).


```{r}
glimpse(flights)
glimpse(planes)

flights %>% 
  
  # left join table flights with planes, that has all the info on airplanes 
  left_join(planes, by="tailnum") %>% 
  
  # for a single variable, count() is the same as group_by() and then summarise()
  # if we add `sort=TRUE`, we get the table sorted in descending order (max to min)
  count(tailnum, sort=TRUE)
```


# For the plane with the greatest number of flights and that had more than 50 seats, please create a table where it flew to during 2013.

```{r}
which_plane <- flights %>% 
  left_join(planes, by="tailnum") %>% 
  filter(seats >= 50) %>% 
  count(tailnum, sort = TRUE) %>% 
  select(tailnum) %>% 
  
  # slice(1) will pull the first, highest number of flights
  slice(1) %>% 
  
  # pull() creates a vector; if we dont use pull() we get a dataframe
  pull()

which_plane


flights %>% 
  filter(tailnum == which_plane) %>% 
  count(origin, dest, sort=TRUE)

```



Problem 4: The nycflights13 package includes a table (weather) that describes the weather during 2013. Use that table to answer the following questions:

# What is the distribution of temperature (`temp`) in July 2013? Identify any important outliers in terms of the `wind_speed`
variable.

```{r}
weather %>% 
  filter(month == 7) %>% 
  ggplot() +
  aes(x= temp)+
  geom_histogram()

# looking at all months
weather %>% 
 # filter(month == 7) %>% 
  ggplot() +
  aes(x= temp)+
  geom_histogram() +
  facet_wrap(~month, nrow=4)+
  theme_light()

# facet_wrap() for all months' wind_speed
weather %>% 
  ggplot() +
  aes(x= wind_speed)+
  geom_histogram()+
  
  facet_wrap(~month, 
             ncol = 3, # to get 3 columns
             scales = "free")+ # easy to check outliers, by having ggplot automatically adjust scales on axes
  
  theme_light()+
  theme(text=element_text(size=12, family="Montserrat"))

```

# Weather correlations

-   What is the relationship between `dewp` and `humid`?
-   What is the relationship between `precip` and `visib`?

```{r}
library(GGally) # package to plot scatterplot- correlation matrix
glimpse(weather)


weather %>% 
  select(dewp, humid, precip, visib, temp) %>% 
  ggpairs()+
  theme_bw()
```



Problem 5: Use the flights and planes tables to answer the following questions:

#  How many planes have a missing date of manufacture?
```{r}
# to get the dataframe
planes %>% 
  filter(is.na(year))

# to get counts
planes %>% 
  count(is.na(year))
```

# What are the five most common manufacturers?

(Hint: you may need to use case_when() to recode the manufacturer name and collapse rare vendors into a category called Other.)

```{r}
planes %>% 
  count(manufacturer, sort = TRUE)

# recode the manufacturer name and collapse rare vendors into a category called `Other`)

planes <- planes %>% 
  mutate(recode_manufacturer  = case_when( # case_when() is an elegant, multiple ifelse()
    manufacturer %in% c("BOEING") ~ "Boeing",
    manufacturer %in% c("AIRBUS INDUSTRIE", "AIRBUS") ~ "Airbus",
    manufacturer %in% c("EMBRAER") ~ "Embraer",
    manufacturer %in% c("MCDONNELL DOUGLAS", "MCDONNELL DOUGLAS AIRCRAFT CO", "MCDONNELL DOUGLAS CORPORATION" ) ~ "McDonnell Douglas",
    TRUE ~ "Other" # everything else will be "Other"
  )) 

```



#   Has the distribution of manufacturer changed over time as reflected by the airplanes flying from NYC in 2013? 

```{r}
flights %>% 
  left_join(planes, by = "tailnum") %>% 
  filter(!is.na(recode_manufacturer)) %>% 
  group_by(month, recode_manufacturer) %>% 
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot()+
  aes(x=factor(month), 
      y = prop, 
      group = recode_manufacturer, 
      colour = recode_manufacturer)+
  geom_line()+
  theme_light()+
  theme(text=element_text(size=12, family="Montserrat"))+
  scale_y_continuous(labels = scales::percent)+
  NULL
  
```



Problem 6: Use the flights and planes tables to answer the following questions:

# What is the oldest plane (specified by tailnum) that flew from NYC in 2013?

```{r}
flights %>% 
  left_join(planes, by = "tailnum") %>%
  rename(year_manufactured = year.y) %>% 
  select(tailnum, year_manufactured) %>% 
  distinct(tailnum, .keep_all = TRUE) %>% 
  arrange(year_manufactured)
```


# How many airplanes that flew from New York City are included in the planes table?

```{r}
flights %>% 
  distinct(tailnum) %>%
  # semi_join() return all rows from x with a match in y.
  semi_join(planes)
  
# quicker. how many palnes from our `planes` table appear in `flights`  
planes %>% 
  semi_join(flights, by="tailnum")

  # An inner_join() only keeps observations from x that have a matching key in y.
flights %>% 
  distinct(tailnum) %>%  
  inner_join(planes)



```



Problem 7: Use the nycflights13 to answer the following questions:

# median arrival delay on a month-by-airport basis 

```{r}
flights %>% 
  group_by(month, origin) %>% 
  summarise(mean_arrival_delay = mean(arr_delay, na.rm=TRUE),
            median_arrival_delay = median(arr_delay, na.rm=TRUE)
            )
```


# plot median arrival delay for each month and origin airport.

```{r}
flights %>% 
  group_by(carrier, month, origin) %>% 
  summarise(mean_arrival_delay = mean(arr_delay, na.rm=TRUE),
            median_arrival_delay = median(arr_delay, na.rm=TRUE)
            ) %>% 
  left_join(airlines, by = "carrier") %>% 
  rename(carrier_name = name) %>% 
  ggplot()+
  aes(x= factor(month),
      y = median_arrival_delay,
      colour = origin, 
      group = origin) +
  geom_line()+
  facet_wrap(~carrier_name, scales = 'free', ncol=3)+
  theme_light() +
    theme(
    text=element_text(size=10, family="Montserrat"))+
  labs(colour = "Origin Airport")
```



Problem 8: Let's take a closer look at what carriers service the route to San Francisco International (SFO). Join the flights and airlines tables and count which airlines flew the most to SFO. Produce a new dataframe, fly_into_sfo that contains three variables: the name of the airline, e.g., United Air Lines Inc. not UA, the count (number) of times it flew to SFO, and the percent of the trips that that particular airline flew to SFO.

# `fly_into_sfo`

```{r}
fly_into_sfo <- flights %>% 
  left_join(airlines, by = "carrier") %>% 
  filter(dest == "SFO") %>% 
  group_by(name) %>% 
  summarise(count = n()) %>% 
  mutate(percent = round(100*count/sum(count),1))

fly_into_sfo

```


```{r}
# And here is some bonus ggplot code to plot your dataframe
fly_into_sfo %>% 
  
  # sort 'name' of airline by the numbers it times to flew to SFO
  mutate(name = fct_reorder(name, count)) %>% 
  
  ggplot() +
  
  aes(x = count, 
      y = name) +
  
  # a simple bar/column plot
  geom_col() +
  
  # add labels, so each bar shows the % of total flights 
  geom_text(aes(label = percent),
             hjust = 1, 
             colour = "white", 
             size = 5)+
  
  # add labels to help our audience  
  labs(title="United (UA) dominates the NYC to SFO route", 
       subtitle = "as % of total flights in 2013",
       x= "Number of flights",
       y= NULL) +
  
  theme_minimal() + 
  
  # change the theme-- i just googled those , but you can use the ggThemeAssist add-in
  # https://cran.r-project.org/web/packages/ggThemeAssist/index.html
  
  theme(#
    # so title is left-aligned
    plot.title.position = "plot",
    
    # text in axes appears larger        
    axis.text = element_text(size=12),
    
    # title text is bigger
    plot.title = element_text(size=18),
    
    text=element_text(size=12, family="Montserrat")
      ) +

  # add one final layer of NULL, so if you comment out any lines
  # you never end up with a hanging `+` that awaits another ggplot layer
  NULL
 
```


# Hollywood Age Gap



```{r, echo=FALSE}
age_gaps <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-02-14/age_gaps.csv')

```





How would you explore this data set? Here are some ideas of tables/ graphs to help you with your analysis.

# How is age_difference distributed? What's the 'typical' age_difference in movies?

```{r}
# select just age_difference and use the skimr package to get summary stats
age_gaps %>% 
  select(age_difference) %>% 
  skimr::skim()

# histogram
age_gaps %>%
  ggplot()+
  aes(x = age_difference)+
  geom_histogram()+
  theme_bw()

# faceted by character 1 gender 
age_gaps %>%
  ggplot()+
  aes(x = age_difference)+
  geom_histogram()+
  theme_bw()+
  facet_wrap(~character_1_gender, ncol = 1)


```


# The half plus seven rule

Large age disparities in relationships carry certain stigmas. One popular rule of thumb is the half-your-age-plus-seven rule. This rule states you should never date anyone under half your age plus seven, establishing a minimum boundary on whom one can date. In order for a dating relationship to be acceptable under this rule, your partner's age must be:

$$\frac{\text{Your age}}{2} + 7 < \text{Partner Age} < (\text{Your age} - 7) * 2$$ 

How frequently does this rule apply in this dataset?

```{r}
age_gaps <- age_gaps %>% 
  mutate(
    lower_limit = actor_1_age/2 + 7,
    half_plus_seven = ifelse(actor_2_age >= lower_limit , 
                             TRUE, 
                             FALSE))

age_gaps %>% 
  count(half_plus_seven) %>% 
  mutate(perc = n/sum(n))


```


# Which movie has the greatest number of love interests? Waht was the median age difference for these movies?

```{r}

age_gaps %>% 
  group_by(movie_name) %>% 
  summarise(n = n(),
            median_age_difference = median(age_difference, na.rm=TRUE)) %>% 
  arrange(desc(n))

```


# Which actors/ actresses have the greatest number of love interests in this dataset?

```{r}
age_gaps %>% 
  count(actor_1_name, sort=TRUE)

age_gaps %>% 
  count(actor_2_name, sort=TRUE)


```


# Is the median age difference staying constant over the years (1935 - 2022)?

```{r}
age_gaps %>% 
  group_by(release_year) %>% 
  summarise(median_difference = median(age_difference, na.rm = TRUE)) %>% 
  ggplot()+
  aes(x=release_year, y = median_difference) + 
  geom_line()+
  geom_smooth(se=FALSE)+
  theme_light()+
  theme(text=element_text(size=12, family="Montserrat"))
            
```



# How frequently does Hollywood depict same-gender love interests?

```{r}
age_gaps <- age_gaps %>% 
  mutate(same_gender = 
           ifelse(character_1_gender == character_2_gender, 
                  TRUE, 
                  FALSE))

age_gaps %>% 
  count(same_gender) %>% 
  mutate(prop = n/sum(n))

age_gaps %>% 
  ggplot()+
  aes(x = release_year, y = age_difference, colour = character_1_gender)+
  geom_point(alpha = 0.6)+
  facet_wrap(~same_gender, ncol=1)+
  theme_bw()+
  theme(text=element_text(size=12, family="Montserrat"))+
  labs(
    colour = "Protagonist's gender"
  )

```

# same-gender over time

```{r}
age_gaps %>% 
  group_by(release_year,same_gender) %>% 
  summarise( n = n()) %>% 
  mutate(prop = n/sum(n)) %>% 
  
  filter(same_gender == FALSE) %>% 
  mutate(same_gender_prop = 1 - prop) %>% 
  ggplot()+
  aes(x=release_year, y = same_gender_prop)+
  
  geom_line()+
  scale_y_continuous(labels = scales::percent)+
  theme_light()


```


# Interactive exploration

```{r}
library(plotly)

plot <- ggplot(age_gaps)+
  geom_point(aes(x=actor_1_age, 
                 y = actor_2_age, 
                 colour = character_1_gender, 
                 
                 # tooltip by default shows all variables used in aes()
                 # to add extra text in tooltip, we use text 
                 text = paste(actor_1_name, "and", actor_2_name, "@",movie_name,"(",release_year,")")), alpha = 0.8)+
  
  # add a line in steelblue that shows lower limit of half plus 7 rule
  geom_line(aes(x = actor_1_age, y = lower_limit),  colour = "steelblue", size = 1.5)+
  theme_light()+
  facet_wrap(~same_gender, ncol=1)


ggplotly(plot)
```

# Most protagonists are men below 40 years

```{r}

base_plot <- age_gaps %>% 
  ggplot()+
  aes(x=actor_1_age, 
      fill = character_1_gender)+
  theme_light()+
  labs(
    title = "Most protagonists are men; majority < 40 years",
    x = "Actors' age",
    fill = "Protagonist's Gender",
    y = NULL
  )+
  theme(text=element_text(size=12, family="Montserrat"))
  
  
base_plot + geom_histogram(alpha = 0.7)
base_plot + geom_density(alpha = 0.6)
base_plot + geom_boxplot(alpha = 0.7)  
base_plot + stat_ecdf(aes(colour = character_1_gender), geom = "step", pad = FALSE) +
  scale_y_continuous(labels = scales::percent)

```

