---
title: "Homework 2"
author: "Your name goes here"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
  pdf_document:
    toc: yes
---


```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false

library(tidyverse)
library(wbstats)
library(skimr)
library(countrycode)
library(here)
library(scales)
library(ggtext)
```

# Data Visualisation - Exploration

Now that you've demonstrated your software is setup, and you have the basics of data manipulation, the goal of this assignment is to practice transforming, visualising, and exploring data.


# Exploring sources of electricity production, CO2 emissions, and GDP per capita.

There are many sources of data on how countries generate their electricity and their CO2 emissions. I would like you to create three graphs:

## 1. A stacked area chart that shows how your own country generated its electricity since 2000.

You will use

`geom_area(colour="grey90", alpha = 0.5, position = "fill")`

## 2. A scatter plot that looks at how CO2 per capita and GDP per capita are related

## 3. A scatter plot that looks at how electricity usage (kWh) per capita/day GDP per capita are related

We will get energy data from the Our World in Data website, and CO2 and GDP per capita emissions from the World Bank, using the `wbstats`package.

```{r}
#| message: false
#| warning: false

# Download electricity data
url <- "https://nyc3.digitaloceanspaces.com/owid-public/data/energy/owid-energy-data.csv"

energy <- read_csv(url) %>% 
  filter(year >= 1990) %>% 
  drop_na(iso_code) %>% 
  select(1:3,
         biofuel = biofuel_electricity,
         coal = coal_electricity,
         gas = gas_electricity,
         hydro = hydro_electricity,
         nuclear = nuclear_electricity,
         oil = oil_electricity,
         other_renewable = other_renewable_exc_biofuel_electricity,
         solar = solar_electricity,
         wind = wind_electricity, 
         electricity_demand,
         electricity_generation,
         net_elec_imports,	# Net electricity imports, measured in terawatt-hours
         energy_per_capita,	# Primary energy consumption per capita, measured in kilowatt-hours	Calculated by Our World in Data based on BP Statistical Review of World Energy and EIA International Energy Data
         energy_per_gdp,	# Energy consumption per unit of GDP. This is measured in kilowatt-hours per 2011 international-$.
         per_capita_electricity, #	Electricity generation per capita, measured in kilowatt-hours
  ) 

# Download data for C02 emissions per capita https://data.worldbank.org/indicator/EN.ATM.CO2E.PC
co2_percap <- wb_data(country = "countries_only", 
                      indicator = "EN.ATM.CO2E.PC", 
                      start_date = 1990, 
                      end_date = 2022,
                      return_wide=FALSE) %>% 
  filter(!is.na(value)) %>% 
  #drop unwanted variables
  select(-c(unit, obs_status, footnote, last_updated)) %>% 
  rename(year = date,
         co2percap = value)


# Download data for GDP per capita  https://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD
gdp_percap <- wb_data(country = "countries_only", 
                      indicator = "NY.GDP.PCAP.PP.KD", 
                      start_date = 1990, 
                      end_date = 2022,
                      return_wide=FALSE) %>% 
  filter(!is.na(value)) %>% 
  #drop unwanted variables
  select(-c(unit, obs_status, footnote, last_updated)) %>% 
  rename(year = date,
         GDPpercap = value)
```

Specific questions:

1.  How would you turn `energy` to long, tidy format?

```{r}
energy %>% 
  filter(country == "Mexico")

energy_long <- energy %>% 
  select(1:12) %>% 
  pivot_longer(cols = 4:12,
               names_to = 'source',
               values_to = 'value') %>% 
  drop_na(value) %>% 
  drop_na(iso_code)
  
  

energy_long %>% 
  filter(country == "Mexico")

energy_long %>% 
  filter(country == "Mexico") %>% 
  ggplot()+
  aes(x= year, y = value, fill = source)+
  geom_area(colour="grey90", alpha = 0.5, position = "fill")

energy_long %>% 
  filter(country == "Mexico") %>% 
  mutate(source = fct_reorder(source, value, max)) %>% 
  ggplot()+
  aes(x= year, y = value, fill = source)+
  geom_area(colour="grey90", alpha = 0.5, position = "fill")


```



2.  You may need to join these data frames
    -   Use `left_join` from `dplyr` to [join the tables](http://r4ds.had.co.nz/relational-data.html)
    -   To complete the merge, you need a unique *key* to match observations between the data frames. Country names may not be consistent among the three dataframes, so please use the 3-digit ISO code for each country
        -   An aside: There is a great package called [`countrycode`](https://github.com/vincentarelbundock/countrycode) that helps solve the problem of inconsistent country names (Is it UK? United Kingdom? Great Britain?). `countrycode()` takes as an input a country's name in a specific format and outputs it using whatever format you specify.
    
    
```{r}
country_names <- energy_long %>% 
  distinct(iso_code, country) %>% 
  select(iso_code, country)

join1 <- gdp_percap %>% 
  left_join(co2_percap, by = c("country", "year")) %>% 
  select(iso_code = iso3c.x, country, year, GDPpercap, co2percap) %>% 
  left_join(country_names, by = "iso_code") %>% 
  rename(country = country.y)


join2 <- join1 %>% 
  left_join(energy, by = c("country", "year")) 


library(countrycode)
join2 <- join2 %>% 
  mutate(
    continent = countrycode(country, 
                            origin = 'country.name',
                            destination = "continent"),
    iso3c = countrycode(country, 
                        origin = 'country.name',
                        destination = "iso3c"),
  ) %>% 
  select(-c(iso_code.x, country.x, iso_code.y))


# Let us downlaod population data from World Bank
# https://data.worldbank.org/indicator/SP.POP.TOTL
# Download population data 
population <- wb_data(country = "countries_only", 
                      indicator = "SP.POP.TOTL", 
                      start_date = 1970, 
                      end_date = 2022,
                      return_wide=FALSE) %>% 
  filter(!is.na(value)) %>% 
  #drop unwanted variables
  select(-c(unit, obs_status, footnote, last_updated))

# join population data with everything else 
join2 <- join2 %>% 
  left_join(population, by=c("iso3c", "year"="date")) %>% 
  rename(population = value,
         country = country.x) %>% 
  select(-c(indicator_id, indicator, iso2c, country.y))
```
    
    
3.  Write a function that takes as input any country's name and returns all three graphs. You can use the `patchwork` package to arrange the three graphs


```{r}
library(viridis)
energy_mix <- function(country) {
  
  energy %>% 
    select(1:2, 4:12) %>% # we will now have 11 columns
    pivot_longer(cols = 3:11,
                 names_to = "source",
                 values_to = "value") %>% 
    
    filter(country == {{country}}, year >= 2000) %>% 
    
    mutate(source = fct_reorder(source, value, max)) %>% 
    drop_na(value) %>% 
    
    
    ggplot()+
    aes(x=year, y = value, fill=source )+
    geom_area(color="grey90", 
              alpha = 0.5,
              position = "fill")+ # so you get to 100%
    theme_bw()+
    scale_fill_viridis(discrete = T,
                       option = "H") +
    scale_y_continuous(labels = scales::percent)+
    labs(title = glue::glue("Energy Production Mix for ", {{country}}),
         x=NULL,
         y=NULL,
         fill= "Source")
}

energy_mix("Germany")

```


A function that plots top n countries for a chosen fuel in a certain year

```{r}
  
energy_long <-   energy %>% 
    select(1:2, 4:12) %>% # we will now have 11 columns
    pivot_longer(cols = 3:11,
                 names_to = "source",
                 values_to = "value") 


top_fuel <- function(.source, .year, .n) {
  
  energy_long %>% 
    
    filter(year == .year,
           source == .source) %>%
    drop_na(value) %>% 
    mutate(country  =  fct_reorder(country, value)) %>%
    slice_max(order_by = value, n = .n) %>%
    ggplot(aes(x = value, y = country)) +
    geom_col() +
    theme_minimal()+
    geom_text(
      aes(label = scales::number(value, accuracy = 0.1), x = value - .25),
      colour = "white",
      size = 4,
      hjust = 1
    ) +
    labs(
      title = paste0("Top ",.n," ", .source, " producing countries ", " in ",.year),
      subtitle = "Terra-Watthours (TWh)",
      x= NULL,
      y = NULL
    )+
    theme(
      plot.title.position = "plot",
  #    plot.title = element_textbox_simple(size=16),
      axis.title.y = element_text(angle = 0, vjust = 0.5,size=14),
      axis.text = element_text(size=12),
      legend.position = "none") +
    NULL
}

top_fuel("gas", 2018, 20)

top_fuel("coal", 2021, 10)

top_fuel("nuclear", 2021, 10)

top_fuel("hydro", 2020, 10)

top_fuel("wind", 2021, 10)

top_fuel("solar", 2021, 15)

```


A function that takes a country as inoput and returns 3 graphs

```{r}
#| message: false
#| warning: false
library(viridis)
library(patchwork)
library(ggrepel)

energy_mix_co2_gdp <- function(country) {
  
  plot1 <- energy %>% 
    select(1:2, 4:12) %>% # we will now have 11 columns
    pivot_longer(cols = 3:11,
                 names_to = "source",
                 values_to = "value") %>% 
    
    filter(country == {{country}}, year >= 2000) %>% 
    drop_na(value) %>% 
    mutate(source = fct_reorder(source, value, max)) %>% 

    
    
    ggplot()+
    aes(x=year, y = value, fill=source )+
    geom_area(color="grey90", 
              alpha = 0.5,
              position = "fill")+ # so you get to 100%
    theme_bw()+
    scale_fill_viridis(discrete = T,
                       option = "H") +
    scale_y_continuous(labels = scales::percent)+
    labs(title = glue::glue("Energy Production Mix for ", {{country}}),
         x=NULL,
         y=NULL,
         fill= "Source")
  
  plot2 <- join2 %>% 
    select(year, country, GDPpercap, co2percap, per_capita_electricity) %>% 
    filter(country == {{country}}, year >= 2000) %>% 
    
    ggplot()+
    aes(x=GDPpercap, y=co2percap, label=year)+
    geom_point()+
    geom_text_repel(size=3)+
    theme_light() +
    scale_x_continuous(labels = scales::dollar)+
    labs(title = "CO2 vs GDP per capita",
         x = "GDP per capita",
         y = "CO2 per capita")
  
  plot3 <- join2 %>% 
    select(year, country, GDPpercap, co2percap, per_capita_electricity) %>% 
    filter(country == {{country}}, year >= 2000) %>% 
    
    ggplot()+
    aes(x=per_capita_electricity/365, y=co2percap, label=year)+
    geom_point()+
    geom_text_repel(size=3)+
    theme_light()+
    labs(title = "CO2 vs per capita electricity consumption",
         x = "Daily Electricity usage (kWh)",
         y = NULL)
  
  
# use patchwork to arrange plots
plot1 / (plot2 + plot3)
}

energy_mix_co2_gdp("Greece")

energy_mix_co2_gdp("India")

energy_mix_co2_gdp("United Kingdom")

energy_mix_co2_gdp("Brazil")

energy_mix_co2_gdp("France")

energy_mix_co2_gdp("China")

energy_mix_co2_gdp("Germany")

energy_mix_co2_gdp("Vietnam")

energy_mix_co2_gdp("Nigeria")

energy_mix_co2_gdp("Argentina")



```
```{r}
mycolours <- c("grey70","tomato")


join2  %>% 
  drop_na(per_capita_electricity) %>% 
  mutate(country = fct_reorder(country, per_capita_electricity),
         myfill = ifelse(country %in% c("China","Denmark","India", "United States", "Italy", "Nigeria"), TRUE, FALSE)) %>% 
  filter(year == 2019, 
         population > 30e6) %>% 
  ggplot()+
  aes(x=per_capita_electricity/365, y = country, fill = myfill,label = number(per_capita_electricity/365,accuracy=0.1))+
  geom_col()+
  facet_wrap(~continent, scales="free")+
  scale_fill_manual(values = mycolours)+
  geom_text(
    colour = "#fafafa",
    size = 3,
    hjust = 1
  )+
  theme_minimal()+
  labs(
    title = "How much energy does the average person use per day?",
    subtitle = "2019 kWh per capita, countries with population > 30m",
    fill = "Energy Source",
    x = NULL,
    y = NULL
  )+
  # ensure title is top-left aligned
  theme(plot.title.position = "plot")+
  theme(legend.position = "none")+
  NULL

```

```{r}
#| message: false
#| warning: false

library(plotly)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgdal)
library(rgeos)
library(patchwork)
library(mapview)
library(tmap)


# per capita kwh vs co2 per cap
co2_kwh_plot <- join2  %>% 
  drop_na(per_capita_electricity) %>% 
  filter(year == 2019, 
         population > 10e6) %>% 
  ggplot()+
  aes(y=per_capita_electricity/365, x = co2percap, colour = continent, 
        size = population^2, 
      label = country)+
  geom_point(alpha = 0.8)+
  labs(
    title = " CO2 per capita vs energy intensity",
    subtitle = "2019 kWh per capita, countries with population > 10m",
    x = "CO2 per capita emmissions",
    y = "kWh per capita per year"
  )+
  coord_flip()+
  theme_minimal()+
  geom_text_repel(size = 4)+
  theme(legend.position = "none")+
  scale_colour_viridis(discrete = T,
                       option = "D")+
  theme(
    plot.title.position = "plot",
    plot.title = element_textbox_simple(size=16),
    axis.text = element_text(size=9),
    legend.position = "none") +
  NULL   


co2_kwh_plot

# Interactive CO2 vs kWh/capita plot with plotly

ggplotly(co2_kwh_plot)

##  World map with CO2 per cap

map <- ne_countries(scale = "medium", returnclass = "sf") %>%
  dplyr::select(name, iso_a3, geometry) %>%
  filter(!name %in% c("Greenland", "Antarctica"))

df <- map %>% 
  left_join(join2 %>% 
              filter(year == 2019), 
            by = c("iso_a3" = "iso3c")) 



base_map <- ggplot(data = df) +
  geom_sf(
    mapping = aes(
      geometry = geometry, #use Natural Earth World boundaries
      fill = co2percap #fill colour = percent vaccinated
    ),
    colour = "#FAFAFA",      # white borders between regions
  )+
  geom_sf_text(aes(label=number(co2percap, accuracy = 0.1)), size=2) + 
  scale_fill_gradientn(
    colours = c('#a6d96a','#d7191c'))+
  labs(title = "2019 CO2 emissions per capita")+
  theme_void()+
  theme(legend.position = "none")+
  NULL

base_map

# Mercator (ew)
map_mercator <- base_map +
  coord_sf(crs = "+proj=merc")

map_mercator


# Robinson projection
map_robinson <- base_map +
  coord_sf(crs = "+proj=robin") 

map_robinson

```



